---
title: "Outcome Framework - GSS shiny house effect "
output:
  html_document:
    css: resources/report.css
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE
  ,message = FALSE
  ,dev.args=list(bg="transparent")
)
```

### Setup, data load and clean

For shiny hosue effect, the following analysis are done for the whole gss population of 4 waves. The change in residence is calculated from the IDI clean address notifiction table. The address notification table defined in IDI's meta data as follow.
"The purpose of [address_notification] is to provide a history of the usual residence address for individuals in the IDI. This would allow a geographic location for the usual residence of an individual to be derived at any given point in time. Available address information is ‘observational’: that is, they record an address on the occasion when a person notifies an agency of a change of address, but these observation points are not generally dates of moving to a location."
In order to reduce the biased in the address notification table (i.e., individuals who appears earlier in the address notification table may have high frequency of change in residence), we fixed the period of 3 year before interview date and after interview date. That is we calculated number of days in a house since the last (latest) move within 3 years of gss interview date. The change of residence was calculated by number of times individual appear with different address register ids within 3 years (before and after) gss interview dates.

```{r setup, include=FALSE}

###Run setup of data and functions

path <- "~/Network-Shares/Datalab-MA/MAA2016-15 Supporting the Social Investment Unit/outcomes_framework/workstream5/rprogs"
setwd(path)

databasename <- "IDI_Clean_20171020"
sandpitname <-"[DL-MAA2016-15]"

source("1_shiny_house_regression_setup.R")

###Source pre-processed dataset
source("2_shiny_house_regression_data.R")


```


### Regression Model - Life satisfaction 10 likert scale for GSS 2014
In this regression model we used demographic variables such as: age, age^2, sex, ethnicity dummy variables (i.e., snz_ethnicity_grp2_nbr represent 'Maori', snz_ethnicity_grp3_nbr represent 'Pacifica', snz_ethnicity_grp4_nbr represent 'Asian', snz_ethnicity_grp5_nbr represent 'Melaa', snz_ethnicity_grp1_nbr represent 'European' and snz_ethnicity_grp6_nbr represent 'Other') , self-explanatory various qualifications indicators, self-reported unemployed indicator, self reported household income (household income is a derived variable, by taking of midpoint of each reported income band and converting to numeric variable. This might not reflect the actual but a proxy of actual household incomes). Using variable 'days_in_house_b4' reflecting the days in a house since last move in a 3 year period before the interview date, and variables 'days_in_house_after' reflecting the days in a house in a period after the interview date. Whereas residence count total, represent change in residence with in 3 before and after interview date from IDI's clean address notification table.
For the shiny house effect we are interested in all significant variables especially in days in house before interview since last move, keeping all other variables at the means levels.
The first model is based on the GSS 2014 survey. The life satisfaction variables is 10 scale likert score and we treat life satisfaction score as a continuous score. This means that if an individual spent more days in a house they were likely to scored low on life satisfaction score. This may indicate shiny house effect.
Moreover, we found the frequent change in residence (resident_count_total) has a negative statistical significant coefficients. This may be refer to stability effect rather shiny house effect. In other words, more you changed your residence more likely you will score low on your life satisfaction score.

In the graphs red asterisk shows significant variables 


```{r reg_2014, echo=FALSE}
reg_10scale_2014 <- lm(gss_pq_feel_life_code ~ age
                         +age2
                         +snz_sex_code
                         #+snz_ethnicity_grp1_nbr (European)
                         +snz_ethnicity_grp2_nbr # Maori
                         +snz_ethnicity_grp3_nbr # Pacifica
                         +snz_ethnicity_grp4_nbr # Asian
                         +snz_ethnicity_grp5_nbr # MELAA
                         +snz_ethnicity_grp6_nbr # Other
                         +sch_qual_ind
                         +higher_sch_qual_ind
                         +tertiary_qual_ind
                         +postgrad_qual_ind
                         +as.factor(unemployed_ind)
                         +log(hh_income_numeric+1)
                         +residence_count_total
                         +log(days_in_house_b4+1)
                         +log(days_in_house_after+1)
                         ,data = gss_var[which(gss_var$gss_id_collection_code=='GSS2014'),]
                         ,weights = gss_pq_person_FinalWgt_nbr)
summary(reg_10scale_2014)

plot_estimates(reg_10scale_2014)
```

### Annex 2 - Regression results showing impact of moving into a new house

Output in the Annex of the Measuring wellbeing impacts of public policy : social housing. 

Life satisfaction scores in all GSS waves were not recorded at 10 likert scale. For standardized score we derived life satisfaction score for all GSS wave at 5 likert scale. The main aim of following regression model were to check whether the relations between the variables remain the same as we had seen in the previous models. Just keep in mind we kept all the above mentioned independent variables with 5 likert score as dependent variable for all waves of GSS.
From the regression we can see all most of independent variables had the same relation as previous model, but number of days before and after interview were not found statistically significant. This means that there was no statistical evidence of shiny house effect on 5 likert Life Satisfaction Score across all GSS waves.
Interestingly, in this model sex is found significant and Female are more likely to be happy in comparison to Male. All dummy ethnicity variables were also found significant except variable snz_ethnicity_grp5_nbr (MELAA), which shows that most ethnicity are less happy when compared to European.
Moreover, all the qualifications indicators were also found significant in comparison to no qualification as a base.



```{r reg_final, echo=FALSE}
reg_5scale_gss <- lm(life_satisfaction_ind ~ age
                         +age2
                         +snz_sex_code
                         #+snz_ethnicity_grp1_nbr (European)
                         +snz_ethnicity_grp2_nbr # Maori
                         +snz_ethnicity_grp3_nbr # Pacifica
                         +snz_ethnicity_grp4_nbr # Asian
                         +snz_ethnicity_grp5_nbr # MELAA
                         +snz_ethnicity_grp6_nbr # Other
                         +sch_qual_ind
                         +higher_sch_qual_ind
                         +tertiary_qual_ind
                         +postgrad_qual_ind
                         +as.factor(unemployed_ind)
                         +log(hh_income_numeric+1)
                         +residence_count_total
                         +log(days_in_house_b4+1)
                         +log(days_in_house_after+1)
                         ,data = gss_var,weights = gss_pq_person_FinalWgt_nbr)
summary(reg_5scale_gss)

plot_estimates(reg_5scale_gss)
```


Logistic Regression Model - Life satistfaction binary scale for 4 waves of GSS
In previous models we treated life satisfaction score (either 10 scale or 5 scale score) as a continuous variables, which might not be right approach. The likert scores are more likely to be ordinal variable than continuous variables. In the following model we used derived binary life satisfaction score in logistic regression model to model shiny house effect.
We kept all the variable used in the first model to see the impact of the shiny house effect on the life satisfaction score.
Overall the relation between life satisfaction binary score and other dependent variables mostly remain the same (in comparison with two previous models). In a nut shell we cannot see any shiny house effect from the model.
Overall education achievement played statistically significant positive roles in life satisfaction score. Ethnicity Maori and Pacifica were statistically score less on their life satisfaction score as compared to base ethnicity of European. Change in residence (residence_count_total) variable were found significant in all models.

```{r gss_logit, echo=FALSE}
gss_glm<-glm(life_satisfaction_bin ~ age
               +age2
               +snz_sex_code
               #+snz_ethnicity_grp1_nbr (European)
               +snz_ethnicity_grp2_nbr # Maori
               +snz_ethnicity_grp3_nbr # Pacifica
               +snz_ethnicity_grp4_nbr # Asian
               +snz_ethnicity_grp5_nbr # MELAA
               +snz_ethnicity_grp6_nbr # Other
               +sch_qual_ind
               +higher_sch_qual_ind
               +tertiary_qual_ind
               +postgrad_qual_ind
               +as.factor(unemployed_ind)
               +log(hh_income_numeric+1)
               +residence_count_total
               +log(days_in_house_b4+1)
               +log(days_in_house_after+1)
             , data = gss_var, family = binomial(logit))
summary(gss_glm)

plot_estimates_glm(gss_glm)
```

 Weighted Logistic Regression Model - Life satistfaction binary scale for 4 waves of GSS
 
For consistency purpose we were calculating the weighted logistic regression to see if gss survey weights had some impact on the model.
From the model we did not see any impact due to weights. The significance of variables were remains the same with or without weight models.
 

```{r gss_logit_wgtd, echo=FALSE}
#To get weighted estimates for logit regression...
survey_design<-svydesign(ids = ~1,data = gss_var, weights=~gss_pq_person_FinalWgt_nbr)

model_wgtd_glm<-svyglm(life_satisfaction_bin ~ age
               +age2
               +snz_sex_code
               #+snz_ethnicity_grp1_nbr (European)
               +snz_ethnicity_grp2_nbr # Maori
               +snz_ethnicity_grp3_nbr # Pacifica
               +snz_ethnicity_grp4_nbr # Asian
               +snz_ethnicity_grp5_nbr # MELAA
               +snz_ethnicity_grp6_nbr # Other
               +sch_qual_ind
               +higher_sch_qual_ind
               +tertiary_qual_ind
               +postgrad_qual_ind
               +as.factor(unemployed_ind)
               +log(hh_income_numeric+1)
               +residence_count_total
               +log(days_in_house_b4+1)
               +log(days_in_house_after+1)
               ,family=quasibinomial(link = 'logit')
               ,design=survey_design)
summary(model_wgtd_glm)
exp(cbind('odd ratio'=coef(model_wgtd_glm), confint(model_wgtd_glm)))

plot_estimates(model_wgtd_glm)
```

Logistic Regression Model - Life satistfaction binary scale, checking for imbalances 

Our derived binary life satisfaction score is highly imbalanced in terms of life satisfaction score. Only 14% of individuals reported not satisfied with their life and this may be referred to our positive class. Even though we are not doing predictions, but still be good to keep close eyes on sensitivity.


```{r gss_logit_imb, echo=FALSE}

table(gss_var$life_satisfaction_bin, useNA = 'always')

prop.table(table(gss_var$life_satisfaction_bin))


### Predict whole dataset
gss_Results <- data.frame(obs = gss_var$life_satisfaction_bin)
gss_Results$prob <- predict(gss_glm, gss_var,type = "response")
gss_Results$pred <- ifelse(gss_Results$prob>0.5,1,0)
gss_Results$Label <- ifelse(gss_Results$obs == "0", 
                              "True Outcome: gss_0", 
                              "True Outcome: gss_1")

### Plot the probability of gss_0
histogram(~prob|Label,
          data = gss_Results,
          layout = c(2, 1),
          nint = 20,
          xlab = "Probability of gss_0",
          type = "count")

### Create the confusion matrix from the gss dataset.
confusionMatrix(data = as.factor(gss_Results$pred), reference = as.factor(gss_Results$obs))

```

Logistic Regression Model - Life satistfaction binary scale, Correcting for imbalances 

In order to balance our targeted class (life satisfaction score with category of zero), we considered only gss population who do not own home. In our gss population we had a variable "housing_groups" and sub setting by the those individuals who were found in those categories: "Accommodation Supplement-Renting","IRR","Non Subsidised-Renting" or "Other Social Housing".

```{r gss_logit_rented, echo=FALSE}
#subsetting by renting groups only....
table(gss_var$housing_groups)

selected<-c("Accommodation Supplement-Renting","IRR","Non Subsidised-Renting" ,"Other Social Housing")

gss_var_rent<-gss_var[gss_var$housing_groups %in% selected,]

dim(gss_var_rent)

table(gss_var_rent$life_satisfaction_bin, useNA = 'always')

prop.table(table(gss_var_rent$life_satisfaction_bin))

model_glm<-glm(life_satisfaction_bin ~ age
               +age2
               +snz_sex_code
               #+snz_ethnicity_grp1_nbr (European)
               +snz_ethnicity_grp2_nbr # Maori
               +snz_ethnicity_grp3_nbr # Pacifica
               +snz_ethnicity_grp4_nbr # Asian
               +snz_ethnicity_grp5_nbr # MELAA
               +snz_ethnicity_grp6_nbr # Other
               +sch_qual_ind
               +higher_sch_qual_ind
               +tertiary_qual_ind
               +postgrad_qual_ind
               +as.factor(unemployed_ind)
               +log(hh_income_numeric+1)
               +residence_count_total
               +log(days_in_house_b4+1)
               +log(days_in_house_after+1)
               ,data = gss_var_rent, family = binomial(logit))
summary(model_glm)

plot_estimates(model_wgtd_glm)

#tidy from the broom for model interpretation....
library(broom)
coefficient<-tidy(model_glm)[ ,c(1:2,4)]

coefficient$estimate<-exp(coefficient$estimate)

coefficient

### Predict whole dataset
gss_Results <- data.frame(obs = gss_var_rent$life_satisfaction_bin)
gss_Results$prob <- predict(model_glm, gss_var_rent,type = "response")
gss_Results$pred <- ifelse(gss_Results$prob>0.5,1,0)
gss_Results$Label <- ifelse(gss_Results$obs == "0", 
                            "True Outcome: gss_0", 
                            "True Outcome: gss_1")

### Plot the probability of gss_0
histogram(~prob|Label,
          data = gss_Results,
          layout = c(2, 1),
          nint = 20,
          xlab = "Probability of gss_0",
          type = "count")


### Create the confusion matrix from the gss dataset.
confusionMatrix(data = as.factor(gss_Results$pred), reference = as.factor(gss_Results$obs))

```

 Regressions - before the interview

Using dummy variables 'b4_***' reflecting the change (or not) of house in a period before the interview date (0-6 months, 6-12 months, 12-24 months, more than 24 months)


```{r regb4_1, echo = FALSE}

Reg_b4_1 <- lm(gss_pq_feel_life_code ~ b4_2Y_3Y
                                   + b4_1Y_2Y
                                   + b4_6M_12M
                                   + b4_6M
            + age
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_b4_1)
plot_estimates(Reg_b4_1)

```

Dummy variables not significant

Adding residence count and months in house before the interview date

Both are significant

```{r regb4_2, echo = FALSE}

Reg_b4_2 <- lm(gss_pq_feel_life_code ~ residence_count_b4 
                                    + months_in_house
                                    + age
                                    + age2
                                    ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_b4_2)
plot_estimates(Reg_b4_2)

```

```{r regb4_3, echo = FALSE}

Reg_b4_3 <- lm(gss_pq_feel_life_code ~ residence_count_b4 
                                    + months_in_house
                                    + b4_2Y_3Y
                                    + b4_1Y_2Y
                                    + b4_6M_12M
                                    + b4_6M
                                    + age
                                    + age2
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_b4_3)
plot_estimates(Reg_b4_3)

```

Hood and house problems highly significant

```{r regb4_4, echo = FALSE}

Reg_b4_4 <- lm(gss_pq_feel_life_code ~ residence_count_b4
           + ct_house_pblms + gss_pq_safe_night_hood_code
           + b4_2Y_3Y
           + b4_1Y_2Y
           + b4_6M_12M
           + b4_6M
           + age
           + age2
           + months_in_house
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_b4_4)
plot_estimates(Reg_b4_4)

```

```{r regb4_5, echo = FALSE}

Reg_b4_5 <- lm(gss_pq_feel_life_code ~ residence_count_b4
           + ct_house_pblms + gss_pq_safe_night_hood_code
           + age
           + age2
           + months_in_house
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_b4_5)
plot_estimates(Reg_b4_5)

```

Regressions - after the interview

Using dummy variables 'after_***' reflecting the change (or not) of house in a period after the interview date (0-6 months, 6-12 months, 12-24 months, more than 24 months)

Change during the year flowing the interview is significant.

```{r Reg_after1, echo = FALSE}
Reg_after1 <- lm(gss_pq_feel_life_code ~  after_2Y_3Y
           + after_1Y_2Y
           + after_6M_12M
           + after_6M
           + age
           + age2
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_after1)
plot_estimates(Reg_after1)
```

```{r Reg_after2, echo = FALSE}
Reg_after2 <- lm(gss_pq_feel_life_code ~ residence_count_after
           + age
           + age2
           + days_in_house_after
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_after2)
plot_estimates(Reg_after2)
```

```{r Reg_after3, echo = FALSE}
Reg_after3 <- lm(gss_pq_feel_life_code ~ residence_count_after
           + age
           + age2
           + days_in_house_after
           + after_2Y_3Y
           + after_1Y_2Y
           + after_6M_12M
           + after_6M
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_after3)
plot_estimates(Reg_after3)
```

```{r Reg_after4, echo = FALSE}
Reg_after4 <- lm(gss_pq_feel_life_code ~ residence_count_after
           + age
           + age2
           + days_in_house_after
           + ct_house_pblms + gss_pq_safe_night_hood_code
           + after_2Y_3Y
           + after_1Y_2Y
           + after_6M_12M
           + after_6M
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_after4)
plot_estimates(Reg_after4)
```

Finally removing dummy variables not significant

```{r Reg_after5, echo = FALSE}
Reg_after5 <- lm(gss_pq_feel_life_code ~ residence_count_after
           + age
           + days_in_house_after
           + ct_house_pblms + gss_pq_safe_night_hood_code
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_after5)
plot_estimates(Reg_after5)
```


 Before and after variables

```{r Reg_1, echo = FALSE}
Reg_1 <- lm(gss_pq_feel_life_code ~  residence_count_b4
          + residence_count_after
           + age
           + days_in_house_after
           + days_in_house_b4
           + ct_house_pblms + gss_pq_safe_night_hood_code
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_1)
plot_estimates(Reg_1)
```

Accommodation supplement significant

```{r Reg_2, echo = FALSE}
Reg_2 <- lm(gss_pq_feel_life_code ~  residence_count_b4
          + residence_count_after
           + age
          + days_in_house_after
           + days_in_house_b4
           + ct_house_pblms + gss_pq_safe_night_hood_code
          + acc_sup_ind
           ,data = gss_var,weights =gss_pq_person_FinalWgt_nbr)
summary(Reg_2)
plot_estimates(Reg_2)
```

